### 2019-04-24 workout_probabilities.Rmd
god this is tough to write!

---
title: "workout_probabilities"
author: "Brian Mannakee"
date: "4/22/2019"
output:
  pdf_document: default
  word_document: default
  html_document: default
---

# Intro

Variant callers are statistical classifiers.
Usefulness of any classifier is a combination of the ranking performance of the classifier, and the quality of decisions based on the classifier in the real world.
Here we describe an empirical Bayes method which uses very high confidence variant calls from MuTect to create a biologically inspired prior probability of mutation based in the mutation spectrum of the tumor.
We use realistic tumor simulations to show that this method is superior to MuTect as a classifier based on AUROC, while equal or inferior to MuTect based on AUPRC except in situations where the distribution of allele frequencies is extreme and unrealistic.
We then implement a method of false discovery rate control that recovers the superiority of the classifier in the decision context.

# Somatic variant calling base statistical model

At every site in the genome with non-zero coverage, Next Generation Sequencing (NGS) produces a vector $\mathbf{x}  = (\{b_i\},\{q_i\}), i = 1\dots D$ of base calls and their associated quality scores, where $D$ is total read depth.
The goal is to use $\mathbf{x}$ to select between competing hypotheses;
$$
\begin{array}{l}
\mathbf{H_0}:\quad \textrm{Alt allele} = m;\quad\nu = 0\\
\mathbf{H_1}:\quad \textrm{Alt allele} = m;\quad\nu = \hat{f},
\end{array}
$$
where $\nu$ is the variant allele frequency, $\hat{f}$ is the maximum likelihood estimate of $\nu$ given data $\mathbf{x}$, i.e. the ratio of the count of variant reads and total read depth, and $m$ is any of the 3 possible alternative non-reference bases.
For a given read with base $b_i$ and q-score $q_i$, the density function under a particular hypothesis is defined as

$$
  \textrm{f}_{\nu,m}(b_i,q_i) = \left\{
    \begin{array}{cr}
      \nu \frac{10^{-q_i/10}}{3} + (1-\nu)(1-10^{-q_i/10}) & b_i = \textrm{reference}\\
      \nu(1-10^{-q_i/10}) + (1-\nu) \frac{10^{-q_i/10}}{3} & b_i = m\\
      \frac{10^{-q_i/10}}{3} & otherwise.
    \end{array}
    \right.
$$

The likelihood under the hypothesis is then $\mathcal{L}_{\nu,m}(\mathbf{x}) = \prod_{i=1}^{D} \textrm{f}_{\nu,m}(x_i)$.
MuTect reports the log likelihood ratio $log(\mathcal{L}_{\nu=\hat{f},m}(\mathbf{x})/\mathcal{L}_{\nu=0,m}(\mathbf{x}))$ as either TLOD or t_lod_fstar depending on the version.
By fixing the threshold posterior odds at two, the site-specific mutation probability a constant $\mathrm{p}(M)= \mu = 3\mathrm{e}{-6}$, and $\mathrm{p}(m \mid M)$ the prior probability of mutation to specific allele $m$ constant $\mathrm{p}(m \mid M) = \mu/3 = 1\mathrm{e}{-6}$, they derive a TLOD threshold of 6.3 for classifying a site as a somatic variant.
Here we examine the effect of the assumption of a constant prior probability of mutation.

# Site-specific prior probability of mutation

While variant calling algorithms typically assume a constant probability of mutation at every site in the genome, work by Alexandrov and others show that the random mutation generating process actually varies from site to site in a nucleotide context specific manner.
We develop a model of the prior probability of mutation to allele $m$ conditional on the observed genomic context $\mathrm{p}(m,M \mid C)$, and demonstrate an empirical Bayes method for computing this probability from MuTect output.
The prior probability $\mathrm{p}(m,M \mid C)$ can be decomposed as
$$
\mathrm{p}(m,M \mid C) = \mathrm{p}(m \mid C) \mathrm{p}(M \mid C) = \mathrm{p}(m \mid C) \mathrm{p}(C \mid M)\frac{p(M)}{p(C)}
$$
since the probability of a mutation at a site and the probability that it is to allele $m$ are independent conditional on the context.
Here $\mathrm{p}(M) = \mu$ as above, and the empirical distribution of contexts $\mathrm{p}(C)$ is the fraction of the genome made up of each context.
We model $\mathrm{p}(C \mid M)$ as a multinomial distribution with parameter $\boldsymbol{\pi} = \{\pi_i\}, i = 1\dots96$.
Mutations are drawn from this multinomial distribution such that $\mathrm{p}(C = i \mid M) = \pi_i$.
The final quantity $\mathrm{p}(m \mid C)$, the probability of mutation to $m$ given a particular three letter context, is a function of $\boldsymbol{\pi}$. 
We are left to estimate only the vector of probabilities $\boldsymbol{\pi}$.

\[
\begin{aligned}
C \mid M,\boldsymbol{\pi} & \sim \textrm{Multinomial}(\boldsymbol{\pi})  \\
\boldsymbol{\pi} \mid \boldsymbol{\alpha} &\sim \textrm{Dirichlet}(\boldsymbol{\alpha}).
\end{aligned}
\]

The posterior distribution of $\boldsymbol{\pi}$ is $\boldsymbol{\pi} \mid C,\boldsymbol{\alpha} \sim \textrm{Dirichlet}(\mathbf{C} + \boldsymbol{\alpha})$, where $\mathbf{C} = (C_1,\dots,C_{96})$ are the counts of mutations present in the tumor for each of the 96 contexts.